Transient Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of Complex Rotators from TESS

### 2023-08-30 paper/ms.tex
Update ms (JNW comments) - CR evolution
Co-authored-by: Joshua Winn <jnwinn@princeton.edu>

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.96\textwidth]{f2.pdf}
		\vspace{-0.3cm}
		\caption{
      {\bf Validation plots used to  label CRs}.  The complete figure
      set, with one image per sector for each of the \nallcands\
      objects in Table~\ref{tab:thetable} is available in the online
      journal. {\bf For internal collaboration review:
      \url{https://www.dropbox.com/sh/khtwk5a6z0zgrif/AABRG21Ire7VS4BUnaD1hOR8a?dl=0}}.
      Panels are as follows.
			{\it a)}: Phase-folded light curve; gray points are raw 2-minute
			data and black points are binned to 200 points per cycle.
      The adopted period is given in the lower-right corner.
			{\it b)}: Phase-dispersion minimization (PDM) periodogram.
			Dotted lines show up to the 10$^{\rm th}$ harmonic and
			subharmonic.
      {\it c)}: DSS finder chart, with 21$''$ and 42$''$ radius
      circles for scale.  One TESS pixel has a full side length of
      21$''$.
			{\it d)}: Cleaned light curve, binned to 20-minute cadence, in
			Barycentric TESS Julian Date (BTJD).
			{\it e)}: Phase-folded light curve, binned to 100 points per
			cycle.  The gray line denotes the spline-fit to the
			wrapped phase-folded light curve, and small gray triangles
			denote automatically identified local minima.
			{\it f)}: Phase-folded light curve at twice the peak period.
			{\it g)}: Phase-folded light curve at half the peak period.
			{\it h)}: Phase-folded time-series within the ``background''
			aperture defined in the SPOC light curves.
			{\it i)}: Phase-folded flux-weighted centroid in the column
			direction.
			{\it j)}: Phase-folded flux-weighted centroid in the row
			direction.
      {\it k)}: Gaia DR2 color--absolute magnitude diagram. The gray
      background denotes stars within 100\,pc from
      \citet{2021A&A...649A...6G}.
      {\it l)}: Information from Gaia DR2, TIC8, and the automated
      dip-counting search pipeline.  ``Neighbors'', abbreviated
      ``nbhr'', are listed within apparent distances of 2 TESS pixels
      if $\Delta T$$<$2.5.
      {\it m)}: BANYAN-$\Sigma$ v1.2 association probabilities,
      calculated using positions, proper motions, and the parallax.
		}
		\label{fig:vet}
	\end{center}
\end{figure*}

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=\textwidth]{f4.pdf}
		\vspace{-0.6cm}
		\caption{
			{\bf CRs keep their periods but change their shapes.}
			Out of the \ngoods\ CRs in Figure~\ref{fig:cqvs}, 32 had
			120-second cadence TESS data available for a baseline of at
			least two years; the 27 brightest are shown here due to space
			constraints.  Each panel shows one sector of TESS data, and is
			phased to its deepest minimum in flux.  Each panel's title shows
			the TIC identifier and approximate period in hours.  Text insets
			show the TESS sector numbers, which generally span two years, or
			at least 1{,}000 cycles.  The vertical scale is fixed across
			sectors to clarify shape changes.  Gray circles are raw 2-minute
			data; colored circles bin to 300 points per cycle. 
		}
		\label{fig:evoln}
	\end{center}
\end{figure*}

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.98\textwidth]{f5.pdf}
		\vspace{-0.3cm}
		\caption{
      {\bf LP 12-502 (TIC~402980664) light curve}, where each time
      segment represents one TESS orbit.  Data were acquired in Sectors
      18-19, 25-26, 53, and 57-58.  Flares are drawn in gray.  The
      light curve is binned to 15-minute intervals so that there are
      96 points per day, and each point is connected by a line.  Data
      gaps are not interpolated; if data are missing, nothing
      is plotted.  The red vertical lines highlight apparently
      instantaneous state changes in the shape of the dip pattern.  
		}
		\label{fig:lplc}
	\end{center}
\end{figure*}

\subsection{Evolution over two year baseline}

Figure~\ref{fig:evoln} shows ``before'' and ``after'' views of 27 CRs
for which TESS 120-second cadence observations were available at least
two years apart.  Such a baseline was available for 32 of the
confirmed \ngoods\ CRs in our catalog; for plotting purposes we show
the brightest 27.  We have phased each sector to its own local minimum
because for most of the sources we do not know the period at the
precision necessary to be able to accurately propagate an ephemeris
over two years.  The achievable period precision, $\sigma_P$, can be
estimated as
\begin{equation}
  \sigma_{P} = \frac { \sigma_{\phi} P } { N_{\rm baseline} },
  \label{eq:periodprecision}
\end{equation}
for $N_{\rm baseline}$ the number of cycles in the observed baseline
and $\sigma_{\phi}$ the phase precision with which any one feature
(e.g.~a dip, or the overall shape of the sinusoidal envelope) can be
tracked.  Assuming $\sigma_\phi$$\approx$$0.02$ and a 20-day baseline
over a single TESS sector yields
$\sigma_{P}$$\approx$$0.25^{+0.38}_{-0.14}$\,minutes for the
population shown in Figure~\ref{fig:evoln}; propagated forward 1{,}000
cycles yields a typical ephemeris uncertainty range of 2-11\,hours.
Measuring the periods for each sector independently, there did not
seem to be any significant ($>$3$\sigma$) period changes.  This
implies an absolute period stability of $\lesssim$0.1\% over the
two-year baseline.

A few objects in Figure~\ref{fig:evoln}
show clear signs of
the CR phenomenon in one sector, and marginal or non-existent signs
in the other.  While there is subjectivity in this assessment, to our
eyes, cases for which at least one sector would be flagged as
``ambiguous'' include
TIC~368129164 (Sector 23 might be labeled an EB),
TIC~177309964 (Sector 38 would be simply a rotating star),
TIC~404144841 (Sector 38 looks like a rotating star),
TIC~201898222 (Sector 3 looks like a rotating star),
TIC~144486786 (Sector 32 might be an RS CVn),
and
TIC~38820496 (Sector 28 might be an RS CVn).
TIC~193831684, assessed on a single-sector basis, would probably be
labeled an eclipsing binary---in fact, \citet{2021ApJ...912..123J}
already gave this source such a label.  However,
based on the shape evolution between Sectors 13 and 39, it is a
CR.  Based on the fraction of sources overall that ``turned off'',
the observed shape evolution implies that CRs have an on-off duty cycle of
$\approx$75\%.  This type of correction is worth including in
population-level estimates of how intrinsically common CRs are in the
low-mass stellar population \citep[e.g.][]{2022AJ....163..144G}.

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=0.98\textwidth]{f6.pdf}
		\vspace{-0.3cm}
		\caption{
			{\bf Evolution of LP 12-502} ($P$=18.5611\,h) at fixed period and
			epoch over three years.  Each panel shows one phase-folded TESS
			orbit; small text denotes relative cycle number.  There are 200
			binned black points per cycle.  The TESS pointing law dictates
			the large time gaps between Cycles 64-248, 315-1233, and
			1264-1410; larger gaps tend to yield larger shape changes.  The
			dips usually evolve over tens to hundreds of cycles.  However
			cycles 1233-1264 show a dip that ``switched'' from a depth and
			duration of 3\% and 3\,hr to 0.3\% and 1\,hr over less than one
			cycle (cf.~Figure~\ref{fig:lplc}).
		}
		\label{fig:lp}
	\end{center}
\end{figure*}

@@

		\subfloat{
			\includegraphics[width=0.46\textwidth]{f7d.pdf}
			\includegraphics[width=0.46\textwidth]{f7c.pdf}
		}
	\end{center}
	\vspace{-0.4cm}
	\caption{
    {\bf River plots of the LP 12-502 light curve}, showing (clockwise
    from top-left) Sectors 18-19, 25-26, 53, and 58-59.  A
    two-harmonic sinusoid has been subtracted to highlight the sharp
    dips.  In Sectors 25-26 (cycles 248-315), periods are visible at
    the fundamental period of 18.5611\,hr, as well as at faster
    ($\phi$$\approx$0-0.07) and slower ($\phi$$\approx$0.25-0.27)
    relative periods based on the presence of blue dips with distinct
    slopes.  Multiple simultaneous periods are also visible in Sectors
    53 and 58-59.  White chunks denote missing data.
	}
	\label{fig:lpriver0}
\end{figure*}

\subsection{Evolution over adjacent sectors, \& LP 12-502}

A small fraction of our objects were in regions of sky with many
sectors of adjacent TESS data.  Out of this already small sample, LP
12-502 (TIC 402980664; $d$=21\,pc, $J$=9.4, $T$=11.1) stood out due to
the quality and content of its data.  We discuss another interesting
source, TIC~300651846, in Appendix~\ref{app:tic3006}.  In this
section, we first present the LP 12-502 observations and then
highlight their implications.

@@

Figure~\ref{fig:lplc} shows all available data for LP 12-502 from TESS
Sectors 18, 19, 25, 26, 53, 58, and 59, split into successive orbits.
The star was observed at 120-second cadence whenever it was observable
by TESS.  We binned the light curve to 15-minute intervals for visual
clarity, and required at least one (120-second cadence) flux
measurement per bin.  Points more than 2.5$\sigma$ above the median
are drawn in gray, also for visual clarity.  Missing data are not
drawn.  Figure~\ref{fig:lp} then shows the same data, but folded for
each successive TESS orbit.  Finally, Figure~\ref{fig:lpriver0} shows
``river plots'' of the same data, split into similar intervals: the
Sector 18-19 data, 25-26 data, 53 data, and 58-59 data.  The river
plots are subject to one additional processing step: we fitted and
subtracted a maximum-likelihood two-harmonic sinusoid independently
from the Sector 18-19 data, 25-26 data, and 53, 58, and 59 data in
order to accentuate changes in the dip timing and structure.

The average period, determined by measuring the PDM peak period over
each sector independently, was $\langle P \rangle = 18.5560$\,hr.  The
range between the maximum and minimum sector-specific periods was
measured to be about one minute.   However, in detail, a period shift
of $\pm$1\,minute would yield major phase drifts over the baseline;
that time interval corresponds to $\approx$1/1000$^{\rm th}$ of a
period, and we have observed 1500 cycles.  By fine-tuning over a grid
in period, we found $P=18.5611 \pm 0.0001$\,hr, which seems to track
certain features in the LP 12-502 light curve well over the entire
dataset.

What exactly is observed?  For the first 64 cycles, the star shows a
pattern reminiscent of an eclipsing binary, with four obvious local
minima.  We dub these dips $\{ 1, 2, 3, 4 \}$ at phases $\{ -0.28,
-0.08, 0, 0.25 \}$, respectively.  Over cycles 0-64, the depth of dips
1 and 3 remain roughly fixed.  However dip 4 decreases in depth by
about 2\%, while dip 2 increases in depth, by about the same amount
(see Figure~\ref{fig:lp}).  During cycles 48-64, a fifth dip may also
be emerging, in the main ``large'' dip group.

There is then a 6-month (184 cycle) gap to Cycles 248-315, which show
two highly structured dip complexes, plus a small leading dip.  The
single leading dip is present at the same phase as in cycles 0-64, and
is therefore likely to be the same structure.  Along a similar line of
logic, it seems plausible that the first ``dip complex'' during cycles
248-264 represents an evolution of the initial complex seen during
cycles 0-64, though with reduced depth.  During cycles 266-310, an
additional local minimum develops between the two complexes; this
feature is best visualized on the river plots
(Figure~\ref{fig:lpriver0}), and we describe its (shorter) period
below.

The second dip complex during cycles 248-315 shows the most
substructure.  During e.g.~cycles 283-298, this single complex shows
six local minima.  The deepest dip is sharp: it shows a flux excursion
of 3.5\% over about 22 minutes (0.02\,$P$), which is the steepest
slope exhibited anywhere in the LP~12-502 dataset.  After the sharp
dip, there is a roughly exponential fall-off spanning about a quarter
of a period, punctuated by coherent local minima and maxima which in
detail (Figure~\ref{fig:lpriver0}) have slightly longer periods than
the sharp dip.  The sharp leading dip only decreases in depth
following a sudden state-switch at BTJD 2030.7 (cycles 299-315), which
happens during a flare (Figure~\ref{fig:lplc}).  The trailing dips
remain thereafter.

Sectors 53--58 (cycles 1233-1481) are comparatively tame; they showed
only four to six dips per cycle.  Some dips remain stable in depth and
duration over this five month interval.  Other dips grow, like the one
at $\phi = +0.06$ between cycles 1499 and 1481.  Other dips, such as
the one at $\phi = +0.12$ in cycles 1233-1264, disappear entirely.
The most dramatic state switch occurs during cycles 1233-1247, when a
large dip ``switches'' from a depth of 3\% and duration of 3 hours to
a depth of 0.3\%, and a duration of 1 hour.

@@

{\sc State-switches reveal dip independence}---The state-switches seen
in cycles 1233-1247 and 299-315 confirm that dips can disappear in
less than one cycle, a point which has been previously appreciated
\citep{2017AJ....153..152S}.  What is new in these particular changes
is that the morphology changes show that the dips can be {\it
independent} and {\it additive}.  For example, throughout cycles
1233-1264, there are three local minima between phases of 0 and 0.3.
They all have identical ingress times.  The shape change during the
transition implies that the leading dip that ``turned off'' (reduced
its depth and duration), while the trailing two dips remained fixed.
This is visible in Figure~\ref{fig:lp}.  In other words, the
structures producing these dips are independent to the degree that one
can undergo a severe change while the others remain identical.  The
state switches during cycles 248-264 and 299-315 share the same
characteristic: it is always the {\it leading} dip of a complex that
``switches off'', leaving the (fixed-depth) trailing dips in its wake.

{\sc Slow growth; rapid death}---Although there are a few instances in
which we observe dips switch off over less than one cycle, dip growth
seems to happen more slowly.  For instance, the dip that grows between
phase 0 and 0.1 between cycles 260-290 begins to become visible around
BTJD 1993.2, and growths in depth by about 2\% over about six cycles,
to become visible by eye by BTJD 1997.7.  The evolution of this
particular dip is most clear in the river plots.  The evolution of the
latter dip group in cycles 1410-1481 is another example of this slow
mode of dip growth.

@@

{\sc Dip durations}---The shortest dip duration for any of the
individual LP 12-502 dips seems to be $\approx$0.06\,$P$ $\approx$
1.08\,hr.  In comparison, using the stellar radius and mass derived in
Section~\ref{subsec:starprops}, the characteristic timescale $T_{\rm
dur} \equiv R_\star P / (\pi a)$ for the transit of a point-source at
corotation is 1.02$\pm$0.10\,hr.  This means that while some of the LP
12-502 dips are sufficiently long to require structures that are
extended in orbital azimuth, the durations of other dips are
consistent with effective radii for the occulting material $R_{\rm
eff} \ll R_\star$.  This implies $a/R_\star \approx 5.8$ for this
material; the analogous timescale at the stellar surface is about six
times slower.

{\sc Dip periods}---Most of the LP 12-502 dips repeat with a period of
$P=18.5611 \pm 0.0001$\,hr.  However the river plots
(Figure~\ref{fig:lpriver0}) reveal that the light curve has specific
dips with simultaneous but distinct periods.  For instance, in sectors
25-26, the local minimum that develops around cycle 262 has a period
faster than the mean period by $\approx$0.1\%, while some of the
trailing local minima in the main dip complex have periods slower than
the mean period, by $\approx$0.04\%.  In addition to the fundamental
period, we were able to identify at least four distinct periods shown
by specific dips over the full Sectors 18-59 dataset, including
periods at 18.5683, 18.5672, 18.5473, and 18.5145\,hr, with a typical
measurement uncertainty of $\approx$0.0002\,hr.  If each period
corresponds to an individual dust or gas clump, then this implies that
multiple distinct clumps can orbit the star simultaneously, at
marginally different separations.

>>>>

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.96\textwidth]{f2.pdf}
		\vspace{-0.3cm}
		\caption{
      {\bf Validation plots used to classify CRs}.  The complete figure
      set, with one image per sector for each of the \nallcands\
      objects in Table~\ref{tab:thetable} is available in the online
      journal. {\bf For internal collaboration review:
      \url{https://www.dropbox.com/sh/khtwk5a6z0zgrif/AABRG21Ire7VS4BUnaD1hOR8a?dl=0}}.
      Panels are as follows.
			{\it a)}: Phase-folded light curve; gray points are raw 2-minute
			data and black points are binned to 200 points per cycle.
      The adopted period is given in the lower-right corner.
			{\it b)}: Phase-dispersion minimization (PDM) periodogram.
			Dotted lines show up to the 10$^{\rm th}$ harmonic and
			subharmonic.
      {\it c)}: DSS finder chart, with 21$''$ and 42$''$ radius
      circles for scale.  One TESS pixel has a full side length of
      21$''$.
			{\it d)}: Cleaned light curve, binned to 20-minute cadence, in
			Barycentric TESS Julian Date (BTJD).
			{\it e)}: Phase-folded light curve, binned to 100 points per
			cycle.  The gray line denotes the spline-fit to the
			wrapped phase-folded light curve, and small gray triangles
			denote automatically identified local minima.
			{\it f)}: Phase-folded light curve at twice the peak period.
			{\it g)}: Phase-folded light curve at half the peak period.
			{\it h)}: Phase-folded time-series within the ``background''
			aperture defined in the SPOC light curves.
			{\it i)}: Phase-folded flux-weighted centroid in the column
			direction.
			{\it j)}: Phase-folded flux-weighted centroid in the row
			direction.
      {\it k)}: Gaia DR2 color--absolute magnitude diagram. The gray
      background denotes stars within 100\,pc from
      \citet{2021A&A...649A...6G}.
      {\it l)}: Information from Gaia DR2, TIC8, and the automated
      dip-counting search pipeline.  ``Neighbors'', abbreviated
      ``nbhr'', are listed within apparent distances of 2 TESS pixels
      if $\Delta T$$<$2.5.
      {\it m)}: BANYAN-$\Sigma$ v1.2 association probabilities,
      calculated using positions, proper motions, and the parallax.
		}
		\label{fig:vet}
	\end{center}
\end{figure*}

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=\textwidth]{f4.pdf}
		\vspace{-0.6cm}
		\caption{
			{\bf Evolution of CR light curves over two years.}
			Out of the \ngoods\ CRs in Figure~\ref{fig:cqvs}, 32 had
			120-second cadence TESS data available for a baseline of at
			least two years; the 27 brightest are shown here due to space
			constraints.  Each panel shows one sector of TESS data, and is
			phased to its deepest minimum in flux.  Each panel's title shows
			the TIC identifier and period in hours.  Text insets
			show the TESS sector numbers, which generally span two years, or
			at least 1{,}000 cycles.  The vertical scale is fixed across
			sectors to clarify shape changes.  Gray circles are raw 2-minute
			data; colored circles bin to 300 points per cycle. 
		}
		\label{fig:evoln}
	\end{center}
\end{figure*}

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.98\textwidth]{f5.pdf}
		\vspace{-0.3cm}
		\caption{
      {\bf LP 12-502 (TIC~402980664) light curve}, where each time
      segment represents one TESS orbit.  Data were acquired in
      Sectors 18-19, 25-26, 53, and 57-58.  Flares are drawn in gray.
      The light curve is binned to 15-minute intervals so that there
      are 96 points per day, and each point is connected by a line.
      Data gaps have nothing plotted.  The red vertical lines
      highlight apparently instantaneous state changes in the shape of
      the dip pattern.  
		}
		\label{fig:lplc}
	\end{center}
\end{figure*}

\subsection{Evolution over two year baseline}
\label{subsec:twoyear}

Figure~\ref{fig:evoln} shows ``before'' and ``after'' views of 27 CRs
for which TESS 120-second cadence observations were available at least
two years apart.  Such a baseline was available for 32 of the
confirmed \ngoods\ CRs in our catalog; for plotting purposes we show
the brightest 27.  
We have defined $\phi=0$ for each sector to be the time of minimum
light observed in that sector, rather than using a consistent phase
definition across multiple sectors.
This is because for most of the sources we do not know the period at the
precision necessary to be able to accurately propagate an ephemeris
over two years.  The achievable period precision, $\sigma_P$, can be
estimated as
\begin{equation}
  \sigma_{P} = \frac { \sigma_{\phi} P } { N_{\rm baseline} },
  \label{eq:periodprecision}
\end{equation}
for $N_{\rm baseline}$ the number of cycles in the observed baseline
and $\sigma_{\phi}$ the phase precision with which any one feature
(e.g.~a dip, or the overall shape of the sinusoidal envelope) can be
tracked.  Assuming $\sigma_\phi$$\approx$$0.02$ and a 20-day baseline
over a single TESS sector yields
$\sigma_{P}$$\approx$$0.25^{+0.38}_{-0.14}$\,minutes for the
population shown in Figure~\ref{fig:evoln}; propagated forward 1{,}000
cycles yields a typical ephemeris uncertainty range of 2-11\,hours.
Measuring the period independently for each sector did not reveal
evidence for significant ($>$3$\sigma$) changes in period, implying
a period stability of $\lesssim$0.1\% over two years.

A few objects in Figure~\ref{fig:evoln} show the CR phenomenon in one
sector, and only marginal signs or no sign of CR behavior in the other
sector.  In our subjective assessment, cases for which at least one
sector would be flagged as ``ambiguous'' include
TIC~368129164 (Sector 23 might be labeled an EB),
TIC~177309964 (Sector 38 would be simply a rotating star),
TIC~404144841 (Sector 38 looks like a rotating star),
TIC~201898222 (Sector 3 looks like a rotating star),
TIC~144486786 (Sector 32 might be an RS CVn),
and
TIC~38820496 (Sector 28 might be an RS CVn).
TIC~193831684, assessed on a single-sector basis, would probably be
labeled an eclipsing binary---in fact, \citet{2021ApJ...912..123J}
already gave this source such a label.  However, based on the shape
evolution between Sectors 13 and 39, it is a CR.  Based on the
fraction of sources overall that ``turned off'', the observed shape
evolution implies that CRs have an on-off duty cycle of $\approx$75\%.
Correcting for the duty cycle might be important in population-level
estimates of the intrinsic frequency of the CR phenomenon
\citep[e.g.][]{2022AJ....163..144G}.

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=0.98\textwidth]{f6.pdf}
		\vspace{-0.3cm}
		\caption{
			{\bf Evolution of LP 12-502} ($P$=18.5611\,h) at fixed period and
      epoch over three years.  Each panel shows one (averaged) TESS
			orbit; small text denotes relative cycle number.  There are 200
			binned black points per cycle.  The TESS pointing law dictates
			the large time gaps between cycles 64-248, 315-1233, and
			1264-1410; larger gaps tend to yield larger shape changes.  The
			dips usually evolve over tens to hundreds of cycles.  However
			cycles 1233-1264 show a dip that switched from a depth and
			duration of 3\% and 3\,hr to 0.3\% and 1\,hr over less than one
			cycle (cf.~Figure~\ref{fig:lplc}).
		}
		\label{fig:lp}
	\end{center}
\end{figure*}

@@

		\subfloat{
			\includegraphics[width=0.46\textwidth]{f7d.pdf}
			\includegraphics[width=0.46\textwidth]{f7c.pdf}
		}
	\end{center}
	\vspace{-0.4cm}
	\caption{
    {\bf River plots of the LP 12-502 light curve}, showing (clockwise
    from top-left) Sectors 18-19, 25-26, 53, and 58-59.  A
    two-harmonic sinusoid has been subtracted to highlight the sharp
    dips.  
    A fixed period and phase are adopted for all sectors; the dips
    across all observations are bounded by $\phi \in [-0.35,0.35]$.
    In Sectors 25-26 (cycles 248-315), periods are visible at
    the fundamental period of 18.5611\,hr, as well as at faster
    ($\phi$$\approx$0-0.07) and slower ($\phi$$\approx$0.25-0.27)
    relative periods based on the presence of blue dips with distinct
    slopes.  Multiple simultaneous periods are also visible in Sector
    53 (cycles 1234-1263) and Sectors 58-59 (cycles 1411-1479).  White
    chunks denote missing data.
    The state changes noted in red in Figure~\ref{fig:lplc} occur in
    cycles 261, 309, and 1241.
	}
	\label{fig:lpriver0}
\end{figure*}

\subsection{Evolution over consecutive sectors, \& LP 12-502}

A few of our complex rotators were near the TESS continuous viewing
zones (Figure~\ref{fig:catalogscatter}, top right).  Out of this
already small sample, LP 12-502 (TIC 402980664; $d$=21\,pc, $J$=9.4,
$T$=11.1) stood out due to the quality and content of its data.  We
discuss another interesting source, TIC~300651846, in
Appendix~\ref{app:tic3006}.  In this section, we describe the LP
12-502 observations and the possible implications.

@@

Whenever LP 12-502 was located within a TESS sector, it was observed
at 120-second cadence.  Figure~\ref{fig:lplc} shows all the available
data, from Sectors 18, 19, 25, 26, 53, 58, and 59. Vertical offsets
were applied to separate the data from different spacecraft orbit
numbers; there are always two orbits per sector.  We binned the light
curve to 15-minute intervals to facilitate visual inspection.  Points
more than 2.5$\sigma$ above the median are drawn in gray, to prevent
outliers from seizing attention.  Data gaps are not connected by lines
(a common source of confusion in light curve visualization).
Figure~\ref{fig:lp} shows the same data after phase-folding each TESS
spacecraft orbit, assuming $P$=18.5611\,hr and a fixed reference epoch
of BTJD=1791.5367.  Finally, Figure~\ref{fig:lpriver0} shows ``river
plots'' of the same data, split into similar intervals: the Sector
18-19 data, 25-26 data, 53 data, and 58-59 data.  The river plots are
subject to one additional processing step: we fitted and subtracted a
maximum-likelihood two-harmonic sinusoid independently from the Sector
18-19 data, 25-26 data, and 53, 58, and 59 data in order to accentuate
changes in the dip timing and structure.

The average period, determined by measuring the PDM peak period over
each sector independently, was $\langle P \rangle = 18.5560$\,hr.  The
range between the maximum and minimum sector-specific periods was
measured to be about one minute.   However, a period shift of
$\pm$1\,minute leads to large phase drifts over the entire timespan of
observations.  One minute is $\approx$1/1000$^{\rm th}$ of a period,
and we have observed 1500 cycles.  By folding with a fine grid of
trial periods, we found that the choice $P=18.5611 \pm 0.0001$ causes
more of the features in the LP 12-502 light curve to maintain constant
phases over the entire dataset.

We now attempt to describe the complex morphology of the light curve
and its evolution.  For the first 64 cycles, the star shows four
obvious local minima.  We dub these dips $\{ 1, 2, 3, 4 \}$ at phases
$\{ -0.28, -0.08, 0, 0.25 \}$, respectively.  Dips 2 and 3 are part of
the same ``global'' minimum, which otherwise resembles a long eclipse.
Over cycles 0-64, the depth of dips 1 and 3 remain roughly fixed.  Dip
4 decreases in depth by about 2\%, and dip 2 increases in depth by
about the same amount (see Figure~\ref{fig:lp}).  A subtle fifth dip
may also be present at phase +0.08, at the end of the global minimum
that includes dips 2 and 3.

There is then a 6-month (184-cycle) gap to Cycles 248-315, which show
two highly structured dip complexes, plus a small leading dip.  The
leading dip has the same phase (relative to minimum light) as in
cycles 0-64, and therefore seems likely to be due to the same
structure.  Along a similar line of logic, it seems plausible that the
first ``dip complex'' during cycles 248-264 represents an evolution
and reduction in amplitude of dips 2 and 3 that were seen during
cycles 0-64.  During cycles 266-310, an additional local minimum
develops between the two complexes; this feature is best visualized on
the river plots (Figure~\ref{fig:lpriver0}), where it is seen to have
a shorter period than the other dips (as described below).

The second dip complex during cycles 248-315 shows the most
substructure.  During e.g.~cycles 283-298, this single complex shows
six local minima.  The first and deepest dip is sharp: it shows a flux
excursion of 3.5\% over about 22 minutes (0.02\,$P$), which is the
steepest slope exhibited anywhere in the LP~12-502 dataset.  After the
sharp dip, there is a roughly exponential return to the baseline flux
spanning about a quarter of a period, punctuated by coherent local
minima and maxima that the river plot (Figure~\ref{fig:lpriver0})
reveals to have slightly longer periods than the sharp dip.  The sharp
leading dip remains roughly constant in amplitude until a sudden
``state change'' at BTJD 2030.7 (cycle 309) that occurred at the same
time as a flare, and left the trailing dips seemingly unaffected.
This apparent state change, and two others, are marked with red lines
in Figure~\ref{fig:lplc}.

The behavior during Sectors 53--58 (cycles 1233-1481) is comparatively tame; the light curve shows
only four to six dips per cycle.  Some dips remain stable in depth and
duration over this five-month interval.  Other dips grow, like the one
at $\phi = +0.06$ between cycles 1499 and 1481.  Other dips, such as
the one at $\phi = +0.12$ in cycles 1233-1264, disappear entirely.
The most dramatic state change occurs during cycle 1241, when a large
dip switches from a depth of 3\% and a duration of 3 hours to a depth
of 0.3\% and a duration of 1 hour.

@@

{\sc State-changes reveal dip independence}---The state-changes seen
in cycles 1233-1247 and 299-315 confirm that dips can disappear in
less than one cycle. While such behavior was also noted by
\cite{2017AJ....153..152S}, the data presented here show further that
the dips can be {\it independent} and {\it additive}.  For example,
throughout cycles 1233-1264, there are three sharp dips between phases
of 0 and 0.3 with different amplitudes but similar slopes.  During the
transition, the leading dip nearly disappeared while the other two
dips hardly changed; compare the centermost two panels of
Figure~\ref{fig:lp}.  Evidently, the material or process responsible
for one dip can vary independently of the materials or processes
responsible for other dips.  The state changes during cycles 248-264
and 299-315 support the same conclusion, while also hinting that the
{\it leading} dip of a complex is most prone to disappearing, leaving
the trailing dips unchanged in its wake.

{\sc Slow growth; rapid death}---LP 12-502 shows at least three
instances in which dips switch off over less than one cycle; we did
not see any such instances of dips switching on.  Dip growth instead
seems to happen more slowly.  For instance, the dip at phase 0-0.1
between cycles 258-290 begins to become detectable during cycle 258,
and growths in depth by about 2\% over the next eight cycles.  The
evolution of this particular dip is most clear in the river plots.
The evolution of the dip group at phases 0.1-0.3 during cycles
1410-1481 is another example of this slow mode of dip growth.

@@

{\sc Dip durations}---The shortest dip duration for any of the
individual LP 12-502 dips seems to be $\approx$0.06\,$P$ $\approx$
1.08\,hr.  This is very similar to the characteristic timescale of of
transiting small body at the corotation radius,
\begin{equation}
T_{\rm dur} \equiv R_\star P_{\rm rot} / (\pi a) = 1.02\pm 0.10\,{\rm hr},
\end{equation}
where we have inserted the stellar radius and mass derived in
Section~\ref{subsec:starprops}.  Thus, the shortest-duration dips
could be produced by transits of bodies or distributions of material
that are smaller than the star.  The corotation radius corresponds to
$a/R_\star \approx 5.8$, i.e., the transit of a body at the corotation
radius has a duration about six times shorter than a feature on the
stellar photosphere that is carried across the visible hemisphere by
rotation.  On the other hand, some dip durations are sufficiently long
that an explanation involving transits would require structures that
are larger than the star along the direction of orbital motion.

{\sc Dip periods}---Most of the LP 12-502 dips repeat with a period of
$P=18.5611 \pm 0.0001$\,hr.  However the river plots
(Figure~\ref{fig:lpriver0}) reveal that a few dips have detectably
distinct periods.  For instance, in sectors 25-26, the dip that
develops around cycle 262 has a period shorter than the mean period by
$\approx$0.1\%, and some of the trailing local minima in the main dip
complex have periods slower than the mean period by $\approx$0.04\%.
In addition to the fundamental period, we were able to identify at
least four distinct periods shown by specific dips over the full
Sectors 18-59 dataset: 18.5683, 18.5672, 18.5473, and 18.5145\,hr,
with a measurement uncertainty of $\approx$0.0002\,hr. Possibly, the
different periods belong to clumps of dust or prominences of gas at
slightly different orbital distances surrounding the corotation
radius.
